---
title: "echan_ex7.3: Testing the Cointegration versus Correlation Properties between KO and PEP
author: "Austin Haffenden"
date: '`r Sys.Date()`'
output: 
  md_document:
    variant: markdown_github
---

```{r setup, include=FALSE}
```

## Testing the Cointegration versus Correlation Properties between KO and PEP

The cointegration test for KO and PEP is the same as that for PEP and KO in 
Example 7.2. 

The cointegration result should shows that the t-statistic for the augmented 
Dicky-Fuller test is −2.14, larger than the 10 percent critical value of 
−3.038, meaning that there is a less than 90 percent probability that these 
two time series are cointegrated.

We will then further test for correlation between the two time series

This is taken from exercise 7.3 of E.Chans book Quantitative Trading. 

```{r cl_libs}
rm(list = ls())
library(tseries)
library(echanFuncs)
```

```{r data_in}
 
# read a spreadsheet named "KO.xls" into MATLAB. 
ko_in <- read.csv(file.path("data", "KO.csv"))

# the first column (starting from the second row) is the trading days in
# format mm/dd/yyyy.
tday1 <- ko_in$Date

# convert the format into yyyymmdd.
tday1 <- format(as.Date(tday1, "%m/%d/%Y"), "%Y%m%d")

# convert the date strings first into cell arrays and then into numeric format.
tday1 <- as.numeric(as.character(tday1))

# the last column contains the adjusted close prices.
adjcls1 <- ko_in$Adj.Close

# read a spreadsheet named "PEP.xls" into MATLAB. 
pep_in <- read.csv(file.path("data", "PEP.csv"))

# the first column (starting from the second row) is the trading days in 
# format mm/dd/yyyy.
tday2 <- pep_in$Date

# convert the format into yyyymmdd.
tday2 <- format(as.Date(tday2, "%m/%d/%Y"), "%Y%m%d")

# convert the date strings first into cell arrays and then into numeric format.
tday2 <- as.numeric(as.character(tday2))

# the last column contains the adjusted close prices.
adjcls2 <- pep_in$Adj.Close

```

```{r merge}
# find all the days when either KO or PEP has data.
tday <- sort(union(tday1, tday2))
adjcls <- matrix(NaN, length(tday), 2)
#========================================
# tday1
foo <- dplyr::intersect(tday1, tday)

# idx1 is the indices of tday1
idx1 <- match(foo, tday1)

# idx is the indices of tday
idx <- match(foo, tday)

adjcls[idx, 1] <- adjcls1[idx1]

#========================================
# tday2
foo <- dplyr::intersect(tday2, tday)

# idx1 is the indices of tday2
idx2 <- match(foo, tday2)

# idx is the indices of tday
idx <- match(foo, tday)

adjcls[idx, 2] <- adjcls2[idx2]
```

```{r calcs}
#========================================

# days where any one price is missing
baddata <- which(rowSums(is.na(adjcls))>0)

# eliminate days when any return is missing
#ret <- ret[-baddata,]
tday <- tday[-baddata]
adjcls <- adjcls[-baddata,]

vnames <- c('KO', 'PEP')

```

```{r cadf}
# run cointegration check using augmented Dickey-Fuller test
# can't find a direct function to do this so taken from quantstart (2018)

# it is unclear as to which series is the dependent variable and which
#is the independent variable for the regression. Thus we will try both
# and make a choice based on the negativity of the ADF test statistic.

comb1 <- lm(adjcls[,1]~adjcls[,2])
comb2 <- lm(adjcls[,2]~adjcls[,1])

# This will provide us with the intercept and regression coefficient for
# these pairs. We can plot the residuals and visually assess the
# stationarity of the series:
  
plot(comb1$residuals, type="l")

comb1

comb2

tseries::adf.test(comb1$residuals, k=1)
tseries::adf.test(comb2$residuals, k=1)
```

```{r hedge}
results <- lm(I(adjcls[,1])~0+adjcls[, 2])
results

# A hedgeRatio of 1.0114 was found. I.e. KO=1.0114*PEP + z, where z can be 
# interpreted as the spread KO-1.0114*PEP, which is not stationary here.
hedgeRatio <- results$coefficients
hedgeRatio

# plot the residuals
z <- results$residuals
plot(z, type = 'l', xlab = "Date (index)", ylab = "Spread ($)", 
     main = "Spread between KO and PEP")
abline(h = 0)
```

```{r correl}
# A test for correlation.
# Should be 
#   1.0000    0.4849
#   0.4849    1.0000
dailyReturns <- (adjcls-lag1(adjcls))/lag1(adjcls)
corrcoef <- cor(dailyReturns[2:nrow(dailyReturns), ])
corrcoef
```

```{r pvals}
# The two time series are significantly correlated.
# corr.test: 0 indicating no association.
corr_p <- cor.test(dailyReturns[2:nrow(dailyReturns), 1], 
                   dailyReturns[2:nrow(dailyReturns), 2])
corr_p
```



