---
title: "echan_ex3.4: Calculating Sharpe Ratio for Long-Only Versus Market-Neutral Strategies"
author: "Austin Haffenden"
date: '`r Sys.Date()`'
output: 
  md_document:
    variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Example 3.4: Calculating Sharpe Ratio for Long-Only Versus
Let’s calculate the Sharpe ratio of a trivial long-only strategy for IGE: buying and holding a share since the close of November 26, 2001, and selling it at close of November 14, 2007. Assume the average risk-free rate during this period is 4 percent per annum in this example. You can download
the daily prices from Yahoo! Finance, specifying the date range desired,
and store them as an Excel file (not the default comma-separated file),
which you can call IGE.xls. The next steps can be done in either Excel or
MATLAB:


```{r cars}
# make sure previously defined variables are erased.
rm(list = ls())
library(readxl) # to read in the xlsx file
```

```{r ige_data}
# read a spreadsheet named "IGE.xls" into MATLAB.
# remove this directory before publishing ***
dir <- "/home/deckard/Dropbox/Ebooks/finance/beginners/epchan_quantTrading/data"
ige_dat <- read_xlsx(file.path(dir, "IGE.xlsx"))
head(ige_dat)                      

# the first column 
# contains the trading days in format yyyy-mm-dd.
tday <- ige_dat$Date

# convert the format into yyyymmdd.
tday <- format(as.Date(tday, "#Y-#m-#d"), "#Y#m#d")
head(tday)

# convert the date strings first into cell arrays and
# then into numeric format.
tday <- as.numeric(as.character(tday))
head(tday)

# the last column contains the adjusted close prices.
cls <- ige_dat$`Adj Close`
head(cls)

data <- data.frame(tday = tday,
                   cls = cls)
head(data)
# sort tday into ascending order and cls along with it
data <- data[order(tday),]
head(data)

end <- length(data$cls)
end
# 1504

# daily returns
dailyret <- (data$cls[2:end]-data$cls[1:end-1])/data$cls[1:end-1]

# excess daily returns assuming risk-free rate of 4#
# per annum and 252 trading days in a year
excessRet <- dailyret - 0.04/252

# the output should be 0.7893
sharpeRatio <- sqrt(252)*mean(excessRet)/sd(excessRet)
sharpeRatio
# 0.7893175
```
```{r spy}
# Assume this is a continuation of the above code.
# Insert your own code here to retrieve data from
# SPY.xls just as above.
# remove this directory before publishing ***
dir <- "/home/deckard/Dropbox/Ebooks/finance/beginners/epchan_quantTrading/data"
spy_dat <- read_xls(file.path(dir, "SPY.xls"))
head(spy_dat)      

# the first column 
# contains the trading days in format yyyy-mm-dd.
tdaySPY <- spy_dat$Date

# convert the format into yyyymmdd.
tdaySPY <- format(as.Date(tdaySPY, "#Y-#m-#d"), "#Y#m#d")
head(tdaySPY)

# convert the date strings first into cell arrays and
# then into numeric format.
tdaySPY <- as.numeric(as.character(tdaySPY))
head(tdaySPY)

# the last column contains the adjusted close prices.
clsSPY <- spy_dat$`Adj Close`
head(clsSPY)

dataSPY <- data.frame(tday = tdaySPY,
                   cls = clsSPY)
head(dataSPY)
# sort tday into ascending order and cls along with it
dataSPY <- dataSPY[order(tdaySPY),]
head(dataSPY)

end <- length(dataSPY$cls)
end
# 1504

# daily returns
dailyretSPY <- (dataSPY$cls[2:end]-dataSPY$cls[1:end-1])/dataSPY$cls[1:end-1]

# excess daily returns assuming risk-free rate of 4#
# per annum and 252 trading days in a year
excessRetSPY <- dailyretSPY - 0.04/252

# net daily returns
# (divide by 2 because we now have twice as much capital.)
netRet <- (dailyret - dailyretSPY)/2

# the output should be 0.783681
sharpeRatioSPY <- sqrt(252)*mean(excessRetSPY)/sd(excessRetSPY)
sharpeRatioSPY
# 0.1820709
```
```{r 3.5}
# Assume this is a continuation of the above MATLAB
# code.

# cumulative compounded returns
cumret <- cumprod(1+netRet)-1
plot(cumret)


[maxDrawdown maxDrawdownDuration]=...
calculateMaxDD(cumret);
[maxDrawdown maxDrawdownDuration]=...
calculateMaxDD(cumret);
# maximum drawdown. Output should be 0.1053
maxDrawdown
# maximum drawdown duration. Output should be 497.
maxDrawdownDuration

# Notice the code fragment above calls a function “calculateMaxDrawdown,”
# which I display below
```
