---
title: 'e_chan_ex3.7: A Simple Mean-Reverting Model with and without Transaction Costs'
author: "Austin Haffenden"
date: "July 15, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Here is a simple mean-reverting model: Buy the stocks with the worst
previous one-day returns, and short the ones with the best previous oneday 
returns. Despite its utter simplicity, this strategy has had great performance since 1995, ignoring transaction costs (it has a Sharpe ratio of 4.47 in 2006). 

Our objective here is to find out what would happen to its performance in 2006 
if we assume a standard 5-basis-point-per-trade transaction cost. (A trade is defined as a buy or a short, not a round-trip transaction.)

The original version of this example uses data from the S&P 500. As I am 
currently working with cryptocurrencies we will use the ecosystem of stocks 
available on the Kraken exchange. This data is availabe via the Kraken API 
which can be downloaded using my KrakenR and KrakenTools packages. 

This example also uses the retrieveHQuoteHistoricalPrice which selects the 
stocks with a unique set of dates. I'm not sure if this is necessary with the
kraken data
```{r retHQuoteHistPrice}
clear
inputFile=‘Export.txt’
outputFile=‘SPX 20071123’
[mysym, mytday, myop, myhi, mylo, mycl, myvol]=...
textread(inputFile, ‘%s %u %f %f %f %f %u’, ...
‘delimiter’, ‘,’)
# Since the single file consists of many symbols,
# we need to find the unique set of symbols.
stocks=unique(mysym)
# Since the single file consists of many repeating set
# of dates for different symbols, we need to find the
# unique set of dates.
tday=unique(mytday)
op=NaN(length(tday), length(stocks))
hi=NaN(length(tday), length(stocks))
lo=NaN(length(tday), length(stocks))
cl=NaN(length(tday), length(stocks))
vol=NaN(length(tday), length(stocks))
fors=1:length(stocks)
stk=stocks{s}
# find the locations (indices) of the data with
# the current symbol.
idxA=strmatch(stk, mysym, ‘exact’)
 # find the
locations (indices) of the data with
# the current set of dates.
[foo, idxtA, idxtB]=intersect(mytday(idxA), tday)
# Extract the set of prices for the current symbol
cktesting
 # from the downloaded data.
op(idxtB, s)=myop(idxA(idxtA))
hi(idxtB, s)=myhi(idxA(idxtA))
lo(idxtB, s)=mylo(idxA(idxtA))
cl(idxtB, s)=mycl(idxA(idxtA))
vol(idxtB, s)=myvol(idxA(idxtA))
end
save(outputFile, ‘tday’, ‘stocks’, ‘op’, ‘hi’, ...
‘lo’, ‘cl’, ‘vol’)


```
```{r code}
rm(list = ls())

startDate=20060101
endDate=20061231

load('SPX_20071123', 'tday', 'stocks', 'cl')

dailyret=(cl-lag1(cl))./lag1(cl) # daily returns

marketDailyret=smartmean(dailyret, 2) # equal weighted market index return

weights=-(dailyret-repmat(marketDailyret, [1 size(dailyret, 2)]))./repmat(smartsum(isfinite(cl), 2), [1 size(dailyret, 2)]) # weight of a stock is proportional to the negative distance to the market index.

weights(~isfinite(cl) | ~isfinite(lag1(cl)))=0 # those stocks that do not have valid prices or daily returns are excluded.

dailypnl=smartsum(lag1(weights).*dailyret, 2)

dailypnl(tday < startDate | tday > endDate) = [] # remove pnl outside of our dates of interest

sharpe=sqrt(252)*smartmean(dailypnl, 1)/smartstd(dailypnl, 1) # Sharpe ratio should be about 0.25

# daily pnl with transaction costs deducted
onewaytcost=0.0005 # assume 5 basis points

weights(tday < startDate | tday > endDate, :) = [] # remove weights outside of our dates of interest

dailypnlminustcost=dailypnl - smartsum(abs(weights-lag1(weights)), 2).*onewaytcost # transaction costs are only incurred when the weights change

sharpeminustcost=sqrt(252)*smartmean(dailypnlminustcost, 1)/smartstd(dailypnlminustcost, 1) # Sharpe ratio should be about -3.19

```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
