---
title: "Preprocessing scraped IBM data from Yahoo"
author: "Austin Haffenden"
date: '`r Sys.Date()`'
output: 
  md_document:
    variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Preprocess the E Chan Ex3.1 data (yahoo IBM)

Further to the E Chan example we will process this data to:
- convert it to the dplyr::tbl_df format to ease manipulations
- remove any rows that contain dividends (May 08 2017 in this case)
- convert the date to the R date format
- fill the gaps in the data (e.g. weekends) 

```{r clr_libs}
rm(list = ls())
library(dplyr)
```

# Read in our data (or download it using https://aushaff.github.io/2017/07/16/e_chan_ex3.1.html)

```{r read_in}
histPriceFile_df <- read.csv("histPriceFile_df.csv", header = TRUE)
histPriceFile_df <- readRDS("histPriceFile_df.rds")

```

# Convert to table dataframe (dplyr) and standardise the column names (if not already done)
```{r names_and_df}
# for ease of later use convert to dplyr::tbl_df
hpf_tbl_df <- tbl_df(histPriceFile_df)
head(hpf_tbl_df)

# remove strange characters from the column names (if not already done)
colnames(hpf_tbl_df) <- c("Date", "Open", "High", "Low", "Close", "Adj_Close", "Volume")
head(hpf_tbl_df)

```

# Remove the rows containing dividends 
```{r remove_divi}
# to remove any rows containing the word Dividend in Open
hpf_tbl_df <- hpf_tbl_df[!grepl("Dividend", 
                                            hpf_tbl_df$Open),]
```

# Convert the date column to Date format
```{r set_date}

# currently the data is stored as chars - convert all columns except
# for the date to a numeric

# Remove commas from the Volume column
hpf_tbl_df$Volume <- gsub(",", "", hpf_tbl_df$Volume)
  
# convert to numeric (new df)
hiPriFi_df  <- hpf_tbl_df
hiPriFi_df[, -1] <- apply(hpf_tbl_df[,-1], 2, as.numeric)

# convert the char date to a numeric and more straighforward format
num_date <- function(in_date) {
  
  in_month <- strsplit(in_date, " ")[[1]][1]
  out_month <- match(in_month, month.abb)
  
  in_day <- strsplit(in_date, " ")[[1]][2]
  in_year <- strsplit(in_date, " ")[[1]][3]

  out_date <- paste(in_day, out_month, in_year, sep = "/")
  return(out_date)
}

# remove the comma from the Date column
hiPriFi_df$Date <-gsub(",", "", hpf_tbl_df$Date)

# apply the num_date function over all the dates
dates <- unlist(sapply(hiPriFi_df$Date, num_date, 
                simplify = FALSE,
                USE.NAMES = FALSE))

# update the dataframe and convert to a date
hiPriFi_df$Date <- dates
hiPriFi_df$Date <- as.Date(hiPriFi_df$Date, format = "%d/%m/%Y")

# view the structure
str(hiPriFi_df)
```

# Fill in the missing days (e.g weekends)

I have been advised that last value forwards and backwards (50% each way)
is the best method to use for machine learning methods with financial time 
series data. However various methods are available to forward and backfill the 
gaps and the library imputeTS has various functions: na.interpolation, na.locf 
and na.kalman. 

At present I have no solid rationale for choosing one method over another. 
There is plenty of discussion and this thesis was highlighted in one of them:
http://lib.tkk.fi/Diss/2010/isbn9789526034539/isbn9789526034539.pdf

Hopefully I will get time to compare methods as implementations progresses. 

# Fill missing days with last value forwards and backwards (50% split)

```{r missing}
# To add the missing days as NA
# get the max and min date from the dataframe
max_date <- max(hiPriFi_df$Date)
min_date <- min(hiPriFi_df$Date) 

# create a sequence of dates from min to max
all_dates <- data.frame(Date=seq(min_date, max_date, by="days"))

# combine the two dataframes so that all dates are included
# the missing dates now have NA values
all_dates_df <- merge(hiPriFi_df, all_dates, 
                      by.x='Date',by.y='Date',all.x=T,all.y=T)

# convert to tbl_df
all_dates_df <- tbl_df(all_dates_df)

```

The best solution to the fill I can find is to:
- count the number of rows for each gap
- add that value as a column
- fill forwards half
- fill backwards half

```{r fill_TS}
# add a column to the data with the number of gaps

# fill forwards with half the number of gaps

# fill backwards with number of gaps

sub <- all_dates_df[1:20,]

df <- data.frame(Open = sub$Open,
                 Close = sub$Close)
# A tibble: 20 x 7
         Date   Open   High    Low  Close Adj_Close  Volume
       <date>  <dbl>  <dbl>  <dbl>  <dbl>     <dbl>   <dbl>
 1 2017-02-23 182.00 182.50 180.92 181.65    179.89 2282500
 2 2017-02-24 180.21 181.49 179.89 181.35    179.60 3300100
 3 2017-02-25     NA     NA     NA     NA        NA      NA
 4 2017-02-26     NA     NA     NA     NA        NA      NA
 5 2017-02-27 181.19 181.25 179.28 179.40    177.66 3691300
 6 2017-02-28 179.38 180.63 179.35 179.82    178.08 3273400
 7 2017-03-01 180.48 182.55 180.03 181.95    180.19 3005600
 8 2017-03-02 181.88 181.88 180.43 180.53    178.78 2918200
 9 2017-03-03 180.53 181.32 179.76 180.05    178.31 1822000
10 2017-03-04     NA     NA     NA     NA        NA      NA
11 2017-03-05     NA     NA     NA     NA        NA      NA
12 2017-03-06 179.72 180.99 179.57 180.47    178.72 3204400

is.na.rle <- rle(is.na(unlist(sub[,2])))
is.na.rle$values & is.na.rle$lengths >= 1
is.na.rle$values <- is.na.rle$values & is.na.rle$lengths >= 3

https://stackoverflow.com/questions/31053915/fill-na-values-with-the-trailing-row-value-times-a-growth-rate

https://stackoverflow.com/questions/16820072/length-of-trend-panel-data/16820379#16820379
https://stackoverflow.com/questions/16842163/consecutive-nas-in-a-column


```

# Now let us plot the data to have a look at it
```{r ohlc_plot}
all_dates_xts <- xts(all_dates_df[,-1], order.by = all_dates_df[,1])

candleChart(all_dates_xts)

```
