---
title: "echan_box6.1: Finding correlations between returns of different time frames"
author: "Austin Haffenden"
date: '`r Sys.Date()`'
output: 
  md_document:
    variant: markdown_github
---

##Finding correlations between returns of different time frames

```{r cl_libs}
rm(list = ls())
detach(package:echanFuncs, unload = TRUE)
library(echanFuncs)
library(R.matlab)


```

```{r data_in}
list.files("data")
data_in <- readMat("data/inputDataOHLCDaily_20120511.mat")
str(data_in)

syms <- data_in$syms
tday <- data_in$tday
cl <- data_in$cl

idx <- match("TU", unlist(syms))

tday <- tday[, idx];
cl <- cl[, idx];

```

## The correlation tests

```{r corr_tests}

for (lookback in c(1, 5, 10, 25, 60, 120, 250)) {
    for (holddays in c(1, 5, 10, 25, 60, 120, 250)) {
      
        ret_lag <- (cl-backshift(lookback, cl))/backshift(lookback, cl)
        ret_fut <- (fwdshift(holddays, cl)-cl)/cl

        # badDates=any([isnan(ret_lag) isnan(ret_fut)], 2);
        badDates <- ifelse(is.na(ret_lag)|is.na(ret_fut),1,0 )
        # ret_lag(badDates)=[];
        ret_lag <- ret_lag[badDates == 0]
        # ret_fut(badDates)=[];
        ret_fut <- ret_fut[badDates == 0]
        
        if (lookback >= holddays) {
          indepSet <- seq(1, length(ret_lag), holddays)
        } else {
          indepSet <- seq(1, length(ret_lag), lookback)
        }
        
        # ret_lag=ret_lag(indepSet);
        ret_lag <- ret_lag[indepSet]
        # ret_fut=ret_fut(indepSet);
        ret_fut <- ret_fut[indepSet]
             
        # [cc, pval]=corrcoef(ret_lag, ret_fut);
        corr_res <- cor.test(ret_lag, ret_fut)
        cc <- corr_res$estimate
        pval <- corr_res$p.value

        # %         fprintf(1, 'lookback=%3i holddays=%3i cc=%7.4f pval=%6.4f\n', lookback, holddays, cc(1, 2), pval(1, 2));
        # fprintf(1, '%3i\t%3i\t%7.4f\t%6.4f\n',  lookback, holddays, cc(1, 2), pval(1, 2));
        cat(lookback, holddays, cc, pval, "\n")
        
  }
}
```

```{r hurst_vratio}
% Hurst exponent and Variance ratio test
H=genhurst(log(cl), 2);
fprintf(1, 'H2=%f\n', H);

% Variance ratio test from Matlab Econometrics Toolbox
[h,pValue]=vratiotest(log(cl));

fprintf(1, 'h=%i\n', h); % h=1 means rejection of random walk hypothesis, 0 means it is a random walk.
fprintf(1, 'pValue=%f\n', pValue); % pValue is essentially the probability that the null hypothesis (random walk) is true.
```

```{r ...}
lookback=250;
holddays=25;

longs=cl > backshift(lookback, cl)  ;
shorts=cl < backshift(lookback, cl) ;

pos=zeros(length(cl), 1);

for h=0:holddays-1
    long_lag=backshift(h, longs);
    long_lag(isnan(long_lag))=false;
    long_lag=logical(long_lag);
    
    short_lag=backshift(h, shorts);
    short_lag(isnan(short_lag))=false;
    short_lag=logical(short_lag);
    
    pos(long_lag)=pos(long_lag)+1;
    pos(short_lag)=pos(short_lag)-1;
end

ret=(backshift(1, pos).*(cl-backshift(1, cl))./backshift(1, cl))/holddays;

ret(isnan(ret))=0;
idx=find(tday==20090102);
% idx=1;

cumret=cumprod(1+ret(idx:end))-1;

plot(cumret);

fprintf(1, 'Avg Ann Ret=%7.4f Ann Volatility=%7.4f Sharpe ratio=%4.2f \n',252*smartmean(ret(idx:end)), sqrt(252)*smartstd(ret(idx:end)), sqrt(252)*smartmean(ret(idx:end))/smartstd(ret(idx:end)));
fprintf(1, 'APR=%10.4f\n', prod(1+ret(idx:end)).^(252/length(ret(idx:end)))-1);
[maxDD maxDDD]=calculateMaxDD(cumret);
fprintf(1, 'Max DD =%f Max DDD in days=%i\n\n', maxDD, round(maxDDD));
fprintf(1, 'Kelly f=%f\n', mean(ret(idx:end))/std(ret(idx:end))^2);

% Avg Ann Ret= 0.0167 Sharpe ratio=1.04 
% APR=    0.0167
% Max DD =-0.024847 Max DDD in days=343
% Kelly f=64.919535


