---
title: "echan_ex3.1: Scrape IBM data from Yahoo"
author: "Austin Haffenden"
date: '`r Sys.Date()`'
output: 
  md_document:
    variant: markdown_github
---

```{r setup, include=FALSE}
```
## Scrape web pags for financial data 

This is the first in a series of posts that will convert the MATLAB scripts 
from E Chan's book 'Quantative Trading: How to Build Your Own Algorithmic
Trading Business' into R code. 

The following code will be used to retrieve a stock's historical price 
information from Yahoo! Finance

```{r cl_libs, message=FALSE}
rm(list = ls()) # clear the workspace
library(httr)
library(quantmod)
library(XML)
```

```{r read_page}

symbol <- "IBM" # the stock of interest
in_url <- paste0('http://finance.yahoo.com/q/hp?s=',  symbol)

# retrieving the webpage data
histPriceFile_page <-GET(in_url)

# convert to a list
histPriceFile <- readHTMLTable(rawToChar(histPriceFile_page$content), 
                               stringsAsFactors = F)

# extract what we want from the list - the second element
histPriceFile_df <- histPriceFile[[2]]

# view the contents
head(histPriceFile_df)

```
The example from the book actually allocates the data to individual variables
and so just for completeness

```{r allocate}

# standardise the column names
colnames(histPriceFile_df) <- c("Date", "Open", "High", "Low", 
                                "Close", "Adj_Close", "Volume")
head(histPriceFile_df)

# allocate to individual variables
op <- histPriceFile_df$Open
hi <- histPriceFile_df$High
lo <- histPriceFile_df$Low
cl <- histPriceFile_df$Close
vol <- histPriceFile_df$Volume
adjCl <- histPriceFile_df$Adj_Close

```
And that is basically it for the example. More needs to be done before we can 
use the data (preprocessing) but that will be another post as I want to keep 
the e chan examples as stand alone items. 

To finish we will save our dataframe for use another time. This can be done 
in a variety of ways dependent on what you want to do with it. A csv file is
useful for sharing and if you want to look at your data outside of R; an rds file
allows you to read you data back in as an R variable.
```{r save}
write.csv(histPriceFile_df, file = "histPriceFile_df.csv")
saveRDS(histPriceFile_df, "histPriceFile_df.rds")

# histPriceFile_df <- read.csv("histPriceFile_df.csv", header = TRUE)
# histPriceFile_df <- readRDS("histPriceFile_df.rds")

```
